---
title: "CS106B: Programming Abstractions"
subtitle: "Lecture 2: Testing Framework"
author: "Olivia Beyer Bruvik"
date: "Fall 2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Lecture 2 {.tabset}

## Topics
- Testing framework
- Miscellaneous

## CS106B Testing Framework

In main(), write:
```{r, eval = FALSE}
runSimpleTests(SELECTED_TESTS;
```

Write tests as:
```{r, eval = FALSE}
EXPECT_EQUAL(functionBeingTested(input), expectedOutput);
EXPECT_EQUAL(generateLyrics(2), "Da Da ");
EXPECT_EQUAL(generateLyrics(3), "Da Da Dat ");
EXPECT_EQUAL(generateLyrics('3'), X); # What happens? Error: data type?
EXPECT_EQUAL(generateLyrics(8), "Da Da Dat Da Da Dat Da Ya Da ");

```

*check out Resources and Testing guide*

## Miscellaneous
- count starts at 0
- constants have ALL_CAPS
- variable names camelCase
- const int - cannot change variable
- LINE UP VARIABLES: 
  + const int DAT_AMOUNT = 3;
  + const int  YA_AMOUNT = 8;
- no magic numbers (random wo/ context)
- text cases comes at bottom of starter code
